### 2021-04-05 Chapter-1/Chapter-1.tex
update
Updates from Overleaf

\chapter{Introduction} \label{chapter1}
\section{Side-channels}
Side channels are ubiquitous in modern computer systems as sensitive
information can leak through many mechanisms such as power,
electromagnetic radiation, and even sound~\cite{agrawal2002side,kar20178,chari1999towards,217605,genkin2014rsa}.
It is typically hard to fix all side-channel vulnerabilities for real-world software systems. Among them, software side-channel attacks, such as cache attacks, memory page attacks, and controlled-channel attacks, are especially problematic as they do not require physical proximity~\cite{7163052,217543,217589,lee2017inferring,191010,liu2015last}. These attacks arises from shared micro-architectural components in a computer processor. By observing inadvertent interactions between two programs, attackers can infer program execution flows that manipulate secrets and guess secrets such as encryption keys~\cite{Osvik2006,Gullasch:2011:CGB:2006077.2006784,203878,10.1007/978-3-540-45238-6_6}.

@@

Various countermeasures have been proposed to defend against
side channel attacks. Hardware-level solutions, such as reducing shared
resources, adopting oblivious RAM, and using transnational
memory~\cite{203878,217537,shih2017t,Zhang:2015:HDL:2775054.2694372}, need new
hardware features and changes in modern complex computer systems, which are
hard to adopt in reality. Therefore, a promising and universal direction is software countermeasures, detecting and eliminating side channel vulnerabilities from code base. It is due to the following reason: many side-channel attacks originate in two code patterns: data flow from secrets to load addresses and data flow from secrets to branch conditions. We refer to them as secret-dependent data accesses and control flows respectively in the dissertation.

On the contrary, software approaches are
usually easy to implement. The basic idea is to eliminate key-dependent
control-flow transfers and data accesses. By removing the above two kinds of code patterns, developers protect their 
code from a series of side-channel attacks.

However, eliminating side channel vulnerabilities from real-world software code base is not easy. First, some vulnerable implementations perform better. For example, RSA implementations usually adopt the CRT optimization,
which is faster but vulnerable to fault attacks~\cite{aumuller2002fault}. Fixing vulnerabilities can introduce new 
weaknesses. Second, it is hard to detect side channels in reality. While recent work~\cite{203878,217537,Wichelmann:2018:MFF:3274694.3274741,Brotzman19Casym,236338,182946} can detect many side-channel vulnerabilities, they either suffer from expensive computation cost or high false positives and false negatives. Recent work~\cite{203878,217537,Wichelmann:2018:MFF:3274694.3274741,Brotzman19Casym,236338,182946} can detect many side-channel vulnerabilities. For example, DATA~\cite{217537} reports 2,248 potential leakage sites for the RSA implementation in OpenSSL 1.1.0f\@. Further analysis shows 1,510 leaks can be dismissed. But that leaves 460 data-access leaks and 278 control-flow leaks. Many of these vulnerabilities have not been fixed by developers for a variety of reasons. Third, most vulnerabilities pose a negligible risk. 
Although some vulnerabilities result in the key being 
entirely compromised~\cite{184415, aumuller2002fault}, 
many others are less severe in reality. Therefore, we need a proper quantification metric to assess the sensitivity of side-channel vulnerabilities, so a developer can efficiently triage them.

Previous work~\cite{182946,5207642} can identify numerous leakages or even provide an upper bound on the amount of leakage, which is useful to verify that an implementation is secure if it incurs zero leakage.
However, these techniques cannot quantify the severity of a leak because they over approximate the leakage. For example, CacheAudit~\cite{182946} 
reports that the upper-bound leakage of AES-128 exceeds the original key size. Besides, existing side-channel quantification work~\cite{182946,5207642} assumes an attacker runs the target program multiple times with different
input secrets and calculates an ``average'' estimation, which is different from real attack scenarios when the secret that an attacker wants to retrieve is fixed. As a consequence, those results are less useful to assess the severity level of each leakage site.

@@

We try to address the above problem from the following two aspects.
\begin{enumerate}
\item \textbf{Detection.} We propose a fast and precise side-channel detection method. Different from previous trace comparison based methods, our method has fewer false positives and false negatives. Compared with methods based on symbolic execution and abstract interpretation, our methods can work on real-world cryptography libraries. Symbolic execution provides fine-grained information, but it is expensive to compute. Therefore, prior symbolic
execution work~\cite{203878,236338,Brotzman19Casym} either analyzes only
small programs or applies domain knowledge~\cite{203878} to simplify the analysis. We examine the bottleneck of the trace-oriented symbolic execution and optimize it to work for real-world crypto-systems.
\item \textbf{Quantification.} We propose methods to quantify information 
leakage precisely from a single trace attack. We quantify the number of bits that can leak during a real execution and define the amount of leaked information as the cardinality of possible secrets based on an attacker's observation. Before an attack, an adversary has a large but finite input space. Whenever the adversary observes a leakage site, they can eliminate some impossible inputs and reduce the input space's size. In an extreme case, if the input space's size reduces to one, an adversary has determined the input, which means all secret information (e.g., the entire secret key) is
leaked. By counting the number of inputs~\cite{10.1007/11499107_24}, we can quantify the information leakage precisely. We use symbolic execution to generate constraints to model the relation 
between the original sensitive input and an attacker's observations. 
\end{enumerate}

\section{Contribution}
We address above problems from two aspects. First, we propose a novel method that can detect side-channel vulnerabilities in real-world applications automatically fast and precisely. Second, we propose methods that can quantify each side-channel leakages precisely. In addition, we study two kinds of threat models, single-trace attacks, and chosen message attacks. In summary, we make the following contributions:

@@

\begin{itemize}
    \item We designs a symbolic analysis method that can detect secret-dependent control-flow transfers and secret-dependent data access. The key idea is to only symbolically execute those machine instructions that are relevant to vulnerabilities while executing other instructions natively.
    \item We implement the above method in a tool called ~\detect{}. The evaluation results shows that ~\detect{}  is 3 to 100 times faster than existing tools while identifying all vulnerabilities when we run it on the same library.  We do not need to apply domain knowledge to only part of the original program, which simplify the analysis. In addition, ~\detect{} yields new leakages which are not identified by previous tools. We confirm those findings with developers.
\end{itemize}

@@

\section{Dissertation Organization}
The rest of the dissertation is organized as follows. We present relevant backgrounds and related works in Chapter~\ref{chapter2}. Chapter~\ref{chapter3} presents a novel method to detect side-channel vulnerabilities in binary executables. Chapter~\ref{chapter4} presents a method that can quantify each side-channel leakage in a single trace attack. Chapter~\ref{chapter5} extends the approach and quantifies the side-channel leakage in chosen message attacks. Chapter~\ref{chapter6} concludes the thesis.

>>>>

\chapter{Introduction} \label{chapter1}
\section{Side-channels}
Side channels are ubiquitous in modern computer systems as sensitive information can leak through many mechanisms such as power, electromagnetic radiation, and even sound~\cite{agrawal2002side,kar20178,chari1999towards,217605,genkin2014rsa}. It is typically hard to fix side-channel vulnerabilities for real-world software systems. Among them, software side-channel attacks, such as cache attacks, memory page attacks, and controlled-channel attacks, are especially problematic as they do not require physical proximity~\cite{7163052,217543,217589,lee2017inferring,191010,liu2015last}. These attacks arises from shared micro-architectural components in a computer processor. By observing inadvertent interactions between two programs, attackers can infer program execution flows that manipulate secrets and guess secrets such as encryption keys~\cite{Osvik2006,Gullasch:2011:CGB:2006077.2006784,203878,10.1007/978-3-540-45238-6_6}.

@@

Various countermeasures have been proposed to defend against side channel attacks. Hardware-level solutions, such as reducing shared resources, adopting oblivious RAM, and using transnational memory~\cite{203878,217537,shih2017t,Zhang:2015:HDL:2775054.2694372}, need new hardware features and changes in modern complex computer systems, which are hard to adopt in reality. Therefore, a promising and universal direction is software countermeasures, detecting and eliminating side channel vulnerabilities (data flow from secrets to load addresses, and data flow from secrets to branch conditions) from code base.

However, eliminating side channel vulnerabilities from real-world software code base is not easy. First, some vulnerable implementations perform better. For example, RSA implementations usually adopt the CRT optimization, which is faster but vulnerable to fault attacks~\cite{aumuller2002fault}. Fixing vulnerabilities can introduce new weaknesses. Second, it is hard to detect side channels in reality. While recent work~\cite{203878,217537,Wichelmann:2018:MFF:3274694.3274741,Brotzman19Casym,236338,182946} can detect many side-channel vulnerabilities, they either suffer from expensive computation cost or high false positives and false negatives. For example, DATA~\cite{217537} reports 2,248 potential leakage sites for the RSA implementation in OpenSSL 1.1.0f\@. Further analysis shows 1,510 leaks can be dismissed. But that leaves 460 data-access leaks and 278 control-flow leaks. Many of these vulnerabilities have not been fixed by developers for a variety of reasons. Third, most vulnerabilities pose a negligible risk. Although some vulnerabilities result in the key being entirely compromised~\cite{184415, aumuller2002fault}, many others are less severe in reality. Therefore, we need a proper quantification metric to assess the sensitivity of side-channel vulnerabilities, so a developer can efficiently triage them.

@@

We try to address the above problem from the following two aspects.
\begin{enumerate}
    \item \textbf{Detection.} We propose a fast and precise side-channel detection method. Different from previous trace comparison based methods, our method has fewer false positives and false negatives. Compared with methods based on symbolic execution and abstract interpretation, our methods can work on real-world cryptography libraries. Symbolic execution provides fine-grained information, but it is expensive to compute. Therefore, prior symbolic execution work~\cite{203878,236338,Brotzman19Casym} either analyzes only small programs or applies domain knowledge~\cite{203878} to simplify the analysis. We examine the bottleneck of the trace-oriented symbolic execution and optimize it to work for real-world applications.
    \item \textbf{Quantification.} We propose methods to quantify information leakage precisely from a single trace attack. We quantify the number of bits that can leak during a real execution and define the amount of leaked information as the cardinality of possible secrets based on an attacker's observation. Before an attack, an adversary has a large but finite input space. Whenever the adversary observes a leakage site, they can eliminate some impossible inputs and reduce the input space's size. In an extreme case, if the input space's size reduces to one, an adversary has determined the input, which means all secret information (e.g., the entire secret key) is leaked. By counting the number of inputs~\cite{10.1007/11499107_24}, we can quantify the information leakage precisely. We use symbolic execution to generate constraints to model the relation between the original sensitive input and an attacker's observations.
\end{enumerate}

\section{Contribution}
This dissertation makes the following contributions. First, we propose a novel method that can detect side-channel vulnerabilities in real-world applications automatically fast and precisely. Second, we propose methods that can quantify each side-channel leakages precisely. In addition, we study two kinds of threat models, single-trace attacks, and multiple-trace attacks. In summary, we make the following contributions:

@@

\begin{itemize}
    \item We designs a symbolic analysis method that can detect secret-dependent control-flow transfers and secret-dependent data access. The key idea is to only symbolically execute those machine instructions that are relevant to vulnerabilities while executing other instructions natively.
    \item We implement the above method in a tool called \detect{}. The evaluation results shows that \detect{}  is 3 to 100 times faster than existing tools while identifying all vulnerabilities when we run it on the same library.  We do not need to apply domain knowledge to only part of the original program, which simplify the analysis. In addition, \detect{} yields new leakages which are not identified by previous tools. We confirm those findings with developers.
\end{itemize}

@@

\section{Dissertation Organization}
The rest of the dissertation is organized as follows. We present relevant backgrounds and related works in Chapter~\ref{chapter2}. Chapter~\ref{chapter3} presents a novel method to detect side-channel vulnerabilities in binary executables. Chapter~\ref{chapter4} presents a method that can quantify each side-channel leakage in a single trace attack. Chapter~\ref{chapter5} extends the approach and quantifies the side-channel leakage in multiple-trace attack. Chapter~\ref{chapter6} concludes the dissertation.
